---
title: "HR vs LR repo"
author: "Miguel ?ngel Armengol de la Hoz"
date: "June 21, 2018"
output: html_document
toc: TRUE
---

# Libraries

```{r}
library(summarytools)
library(RPostgreSQL)
library(dplyr)
library(sqldf)
```


# Creating useful funtions

```{r}
# function avoiding add the connection data in each iteration
run_query<-function(query){
query_output<-dbGetQuery(con,query)
  return(query_output)
}

# function for reading sql files
getSQL <- function(filepath){
  con = file(filepath, "r")
  sql.string <- ""

  while (TRUE){
    line <- readLines(con, n = 1)

    if ( length(line) == 0 ){
      break
    }

    line <- gsub("\\t", " ", line)

    if(grepl("--",line) == TRUE){
      line <- paste(sub("--","/*",line),"*/")
    }

    sql.string <- paste(sql.string, line)
  }

  close(con)
  return(sql.string)
}
```


##Patients developing AKI during ICU stay

### Data extraction

Tables needed for calculating AKI stage.

#### Chronic patients with AKI receiving rrt

(they already had AKI before ICU admin)

```{r}

chronicAKI<-run_query(getSQL('sql/aki/chronicAKI.sql'))
```

#### Baseline Creatinine

first available value between -12 and +12h from admission [VER 25/01/17]

```{r}
baseline_creat<-run_query(getSQL('sql/aki/baseline_creat.sql'))
```

#### Peak creatinine within 48h

Peak creatinine within first 48h + time from admission to peak creatinine within first 48h

```{r}
peakcreat48h<-run_query(getSQL('sql/aki/peakcreat48h.sql'))
```

#### Peak creatinine within first 7 days

Peak creatinine within first 7 days + time from admission to peak creatinine in 7 days + time from peak creatinine in 7 days to ICU discharge

```{r}
peakcreat7days<-run_query(getSQL('sql/aki/peakcreat7days.sql'))
```



### AKI Dataset creation

```{r}
library(dplyr)
#We need to exclude patients who already had AKI before ICU admin)
AKIdevelopment<-subset(baseline_creat, !(baseline_creat$patientunitstayid %in% chronicAKI$patientunitstayid) )
AKIdevelopment<-left_join(AKIdevelopment, peakcreat7days)
AKIdevelopment<-left_join(AKIdevelopment, peakcreat48h)
AKIdevelopment$position<-NULL
```


### Windsorization

Clean data by means of winsorization, i.e., by shrinking outlying observations to the border of the main part of the data.
The creatinine values have been winsorize replacing the extreme observations using 95% and 5% percentiles.

```{r}
AKIdevelopmentW<-AKIdevelopment
## Winsorization: Replace the extreme observations using 95% and 5% percentiles
winsorize_x = function(x, cut = 0.05){
  cut_point_top <- quantile(x, 1 - cut, na.rm = T)
  cut_point_bottom <- quantile(x, cut, na.rm = T)
  i = which(x >= cut_point_top) 
  x[i] = cut_point_top
  j = which(x <= cut_point_bottom) 
  x[j] = cut_point_bottom
  return(x)
}

AKIdevelopmentW$creat1<-winsorize_x(AKIdevelopmentW$creat1)
AKIdevelopmentW$peakcreat7d<-winsorize_x(AKIdevelopmentW$peakcreat7d)
AKIdevelopmentW$peakcreat48h<-winsorize_x(AKIdevelopmentW$peakcreat48h)

#Now we turn all emptpy values into 0

AKIdevelopmentW$creat1[is.na(AKIdevelopmentW$creat1)]<-0
AKIdevelopmentW$peakcreat7d[is.na(AKIdevelopmentW$peakcreat7d)]<-0
AKIdevelopmentW$peakcreat48h[is.na(AKIdevelopmentW$peakcreat48h)]<-0




```

###List of patients with AKI and offset

```{r}

AKIdevelopmentW['cr7undercreat1']<-AKIdevelopmentW$peakcreat7d/AKIdevelopmentW$creat1
AKIdevelopmentW['cr48hdundercreat1']<-AKIdevelopmentW$peakcreat48h-AKIdevelopmentW$creat1
AKIdevelopmentW['AKIoffset']<-NA
AKIdevelopmentW['AKIstage']<-NA

#nested if else function

i <- function(if_stat, then) {
  if_stat <- lazyeval::expr_text(if_stat)
  then    <- lazyeval::expr_text(then)
  sprintf("ifelse(%s, %s, ", if_stat, then)
}

e <- function(else_ret) {
  else_ret <- lazyeval::expr_text(else_ret)
  else_ret
}

ie <- function(...) {
  args <- list(...)
  
  for (i in 1:(length(args) - 1) ) {
      if (substr(args[[i]], 1, 6) != "ifelse") {
        stop("All but the last argument, need to be i functions.", call. = FALSE)
      }
  }
  if (substr(args[[length(args)]], 1, 6) == "ifelse"){
    stop("Last argument needs to be an e function.", call. = FALSE)
  }
  args$final <- paste(rep(')', length(args) - 1), collapse = '')
  eval_string <- do.call('paste', args)
  eval(parse(text = eval_string))
}

AKIdevelopmentW$AKIstage <- 
  ie(
    i(AKIdevelopmentW$cr7undercreat1>=3,   3),
    i(AKIdevelopmentW$cr7undercreat1>=2,   2),
    i(AKIdevelopmentW$cr7undercreat1>=1.5,   1),
    i(AKIdevelopmentW$cr48hdundercreat1>=0.3,   1),
    e(0)
  )

AKIdevelopmentW$AKIoffset <- 
  ie(
    i(AKIdevelopmentW$cr7undercreat1>=3,AKIdevelopmentW$peakcreat7d_offset),
    i(AKIdevelopmentW$cr7undercreat1>=2,AKIdevelopmentW$peakcreat7d_offset),
    i(AKIdevelopmentW$cr7undercreat1>=1.5,AKIdevelopmentW$peakcreat7d_offset),
    i(AKIdevelopmentW$cr48hdundercreat1>=0.3,   AKIdevelopmentW$peakcreat48h_offset),
    e(0)
  )

library(dplyr)
AKIlist<-AKIdevelopmentW %>%
  select(patientunitstayid,AKIstage
         # ,AKIoffset The nephrologist suggested not to use this variable since we can not infer when AKI was developed from it. Removing it to avoid confusion.
         )
```

# First RRT

Extracts the first rrt from non chronic patients from treatment and the first rrt from intakeoutput separately. 
Then it selects the first rrt between treatment and intakeoutput per patient.

```{r}
# query the data from postgreSQL 
first_rrt <-run_query(getSQL('sql/aki/first_rrt.sql'))
```

# Final List of patients with AKI and offset (with RRT)

Offset is not reliable so we exclude it.

```{r}
# Dr Danziger said the offset is not reliable for telling when the patient exactly developed AKI so he adviced not to use it

AKIlist_final<-left_join(AKIlist,first_rrt)


AKIlist_final$AKIstage <- 
  ie(
    i(!is.na(AKIlist_final$first_rrtoffset),   3),
    e(AKIlist_final$AKIstage)
  )

# The nephrologist suggested not to use this variable since we can not infer when AKI was developed from it. Removing it to avoid confusion.
# AKIlist_final$AKIoffset <- 
#   ie(
#     i(!is.na(AKIlist_final$first_rrtoffset), AKIlist_final$first_rrtoffset),
#     e(AKIlist_final$AKIoffset)
#   )

AKIlist_final$first_rrtoffset<-NULL
AKIlist_final[is.na(AKIlist_final)]<-0
```

# Septic Patients + demographics

We are only interested on the first ICU stay that's why we do position =1
 
```{r}
septic_patients_demograph<-run_query(getSQL('sql/septic_patients_demograph.sql'))
# excluding patients with outliers
septic_patients_demograph$unabridgedactualventdays[septic_patients_demograph$unabridgedactualventdays>60]<-NA

bmi_related<-run_query(getSQL('sql/bmi_related.sql'))
```

# Urine Output Dialysis

Urine Output / dialysis / total per day for first 3 days

```{r}
uo_dialysis<-run_query(getSQL('sql/uo_dialysis.sql'))
# removing outliers
uo_dialysis[uo_dialysis<0]<-NA
```

# Charlson score

The basic objective of this paper is to evaluate an age-comorbidity index in a cohort of patients who were originally enrolled in a prospective study to identify risk factors for peri-operative complications. Two-hundred and twenty-six patients were enrolled in the study. The participants were patients with hypertension or diabetes who underwent elective surgery between 1982 and 1985 and who survived to discharge. Two-hundred and eighteen patients survived until discharge. These patients were followed for at least five years post-operatively. The estimated relative risk of death for each comorbidity rank was 1.4 and for each decade of age was 1.4. When age and comorbidity were modelled as a combined age-comorbidity score, the estimated relative risk for each combined age-comorbidity unit was 1.45. Thus, the estimated relative risk of death from an increase of one in the comorbidity score proved approximately equal to that from an additional decade of age. The combined age-comorbidity score may be useful in some longitudinal studies to estimate relative risk of death from prognostic clinical covariates.
--Charlson M, Szatrowski TP, Peterson J, Gold J. Validation of a combined comorbidity index. J Clin Epidemiol. 1994;47(11):1245-51. PMID: 7722560
--Online calculator:
--http://www.pmidcalc.org/?sid=7722560&newtest=Y

```{r}
charlson_score<-run_query(getSQL("sql/charlson_score.sql"))
```

# Oasis

Oxford Acute Severity of Illness Score (OASIS)
-- This query extracts the Oxford acute severity of illness score in the eICU database.
-- This score is a measure of severity of illness for patients in the ICU.
-- The score is calculated on the first day of each ICU patients' stay.
-- Reference for OASIS:
--    Johnson, Alistair EW, Andrew A. Kramer, and Gari D. Clifford.
--    "A new severity of illness scale using a subset of acute physiology and chronic health evaluation data elements shows comparable predictive accuracy*."
--    Critical care medicine 41, no. 7 (2013): 1711-1718.
 -- Variables used in OASIS:
--  Heart rate, GCS, MAP, Temperature, Respiratory rate, Ventilation status
--  Urine output
--  Elective surgery
--  Pre-ICU in-hospital length of stay
--  Age
 -- Note:
--  The score is calculated for *all* ICU patients, with the assumption that the user will subselect appropriate ICUSTAY_IDs.
--  For example, the score is calculated for neonates, but it is likely inappropriate to actually use the score values for these patients.
 -- TODO:
-- the current query relies only on variables pre-recorded for the APACHE-IV score
-- it may be advisable to use raw values for vital signs instead (HR, MAP, temp, RR)
-- and record min and max values for the first 24h after ICU admission
-- Some missing values in UO could be retrieved by extracting data from intakeoutput table

```{r}
oasis<-run_query(getSQL('sql/oasis.sql'))
```

# SOFA

```{r}
sofa_total_open<-run_query(getSQL('sql/sofa/sofa_total_open.sql'))
```

# Treatment Vasopressor

```{r}
treatment_vasopressor<-run_query(getSQL('sql/treatment_vasopressor.sql'))
```

# Hypotension Time

```{r}
hypotension_time<-run_query(getSQL('sql/hypotension_time/hypotension_time.sql'))
```

# Blood gas acidosis

```{r}
bg_acidosis<-run_query(getSQL('sql/bg_acidosis.sql'))
```

# CKD history on admission

```{r}
adm_chronic_kidney_disease<-run_query(getSQL('sql/adm_chronic_kidney_disease.sql'))
```

# FINAL JOIN

```{r}
# First of all we exclude patients with chronic AKI

hr_dataset<-sqldf("SELECT * 
FROM septic_patients_demograph A
LEFT JOIN chronicAKI B
ON A.patientunitstayid = B.patientunitstayid
WHERE B.patientunitstayid IS NULL ",drv="SQLite")

hr_dataset<-Reduce(function(...) merge(..., all.x=TRUE), list(
 hr_dataset
,bmi_related
,AKIlist_final
,first_rrt
,uo_dialysis
,charlson_score
,oasis
,sofa_total_open
,treatment_vasopressor
,hypotension_time
,bg_acidosis
,adm_chronic_kidney_disease
))

hr_dataset<-unique(hr_dataset)
```

# Summary

```{r}
view(dfSummary(hr_dataset))
```



