---
title: "01_dataset_creation"
author: "Miguel Ángel Armengol"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_notebook:
    code_folding: hide
    number_sections: yes
    theme: flatly
    toc: yes
    toc_float: yes
knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = paste0(substr(inputFile,1,nchar(inputFile)-4)," ",Sys.Date(),'.html')) })
editor_options: 
  markdown: 
    wrap: 72
---

# Environment

```{r message=FALSE, warning=FALSE}
library(summarytools)
library(dplyr)
library(sqldf)
library(tidyverse)
library(nlme)
library(zoo)
library(tableone)
library(readxl)
library(feather)
library(magrittr)
library(MatchIt)
library(kableExtra)

#not in formula

'%!in%' <- function(x,y)!('%in%'(x,y))

# function avoiding add the connection data in each iteration
run_query<-function(query){
query_output<-dbGetQuery(con,query)
  return(query_output)
}

# function for reading sql files from a folder
getSQL <- function(filepath){
  con = file(filepath, "r")
  sql.string <- ""

  while (TRUE){
    line <- readLines(con, n = 1)

    if ( length(line) == 0 ){
      break
    }

    line <- gsub("\\t", " ", line)

    if(grepl("--",line) == TRUE){
      line <- paste(sub("--","/*",line),"*/")
    }

    sql.string <- paste(sql.string, line)
  }

  close(con)
  return(sql.string)
}
```

# Data extraction

## AKI developed during ICU stay calculation

Tables needed for calculating AKI stage.

### Dialysis

```{r}
# query the data from postgreSQL 
dialysis <-run_query(getSQL('sql/aki/dialysis.sql'))
```

### Baseline Creatinine

first available value between -12 and +12h from admission [VER 25/01/17]

```{r}
baseline_creat<-run_query(getSQL('sql/aki/baseline_creat.sql'))
```

### Peak creatinine within 48h

Peak creatinine within first 48h + time from admission to peak
creatinine within first 48h

```{r}
peakcreat48h<-run_query(getSQL('sql/aki/peakcreat48h.sql'))
```

### Peak creatinine within first 7 days

Peak creatinine within first 7 days + time from admission to peak
creatinine in 7 days + time from peak creatinine in 7 days to ICU
discharge

```{r}
peakcreat7days<-run_query(getSQL('sql/aki/peakcreat7days.sql'))
```

### AKI Dataset creation

```{r}
library(dplyr)
#We need to exclude patients who have history_of_esrd )
AKIdevelopment<-subset(baseline_creat, !(baseline_creat$patientunitstayid %in% dialysis$patientunitstayid[dialysis$history_of_esrd_bin==1]) )
AKIdevelopment<-left_join(AKIdevelopment,dialysis%>%select(patientunitstayid,acute_dialysis_bin))
AKIdevelopment<-left_join(AKIdevelopment, peakcreat7days)
AKIdevelopment<-left_join(AKIdevelopment, peakcreat48h)
```

### Windsorization

This function Cleans data by means of winsorization, i.e., by shrinking
outlying observations to the border of the main part of the data. The
creatinine values have been winsorized replacing the extreme
observations using 95% and 5% percentiles.

```{r}
AKIdevelopmentW<-AKIdevelopment
## Winsorization: Replace the extreme observations using 95% and 5% percentiles
winsorize_x = function(x, cut = 0.05){
  cut_point_top <- quantile(x, 1 - cut, na.rm = T)
  cut_point_bottom <- quantile(x, cut, na.rm = T)
  i = which(x >= cut_point_top) 
  x[i] = cut_point_top
  j = which(x <= cut_point_bottom) 
  x[j] = cut_point_bottom
  return(x)
}

AKIdevelopmentW$creat1<-winsorize_x(AKIdevelopmentW$creat1)
AKIdevelopmentW$peakcreat7d<-winsorize_x(AKIdevelopmentW$peakcreat7d)
AKIdevelopmentW$peakcreat48h<-winsorize_x(AKIdevelopmentW$peakcreat48h)

#Now we turn all emptpy values into 0

AKIdevelopmentW$creat1[is.na(AKIdevelopmentW$creat1)]<-0
AKIdevelopmentW$peakcreat7d[is.na(AKIdevelopmentW$peakcreat7d)]<-0
AKIdevelopmentW$peakcreat48h[is.na(AKIdevelopmentW$peakcreat48h)]<-0




```

### List of patients with AKI and offset

```{r}

AKIdevelopmentW['cr7undercreat1']<-AKIdevelopmentW$peakcreat7d/AKIdevelopmentW$creat1
AKIdevelopmentW['cr48hdundercreat1']<-AKIdevelopmentW$peakcreat48h-AKIdevelopmentW$creat1
AKIdevelopmentW['AKIstage']<-NA

#nested if else function

i <- function(if_stat, then) {
  if_stat <- lazyeval::expr_text(if_stat)
  then    <- lazyeval::expr_text(then)
  sprintf("ifelse(%s, %s, ", if_stat, then)
}

e <- function(else_ret) {
  else_ret <- lazyeval::expr_text(else_ret)
  else_ret
}

ie <- function(...) {
  args <- list(...)
  
  for (i in 1:(length(args) - 1) ) {
      if (substr(args[[i]], 1, 6) != "ifelse") {
        stop("All but the last argument, need to be i functions.", call. = FALSE)
      }
  }
  if (substr(args[[length(args)]], 1, 6) == "ifelse"){
    stop("Last argument needs to be an e function.", call. = FALSE)
  }
  args$final <- paste(rep(')', length(args) - 1), collapse = '')
  eval_string <- do.call('paste', args)
  eval(parse(text = eval_string))
}

AKIdevelopmentW$AKIstage <- 
  ie(
    i(AKIdevelopmentW$cr7undercreat1>=3 || AKIdevelopmentW$acute_dialysis_bin==1,   3),
    i(AKIdevelopmentW$cr7undercreat1>=2,   2),
    i(AKIdevelopmentW$cr7undercreat1>=1.5,   1),
    i(AKIdevelopmentW$cr48hdundercreat1>=0.3,   1),
    e(0)
  )

library(dplyr)
AKIlist<-AKIdevelopmentW %>%
  select(patientunitstayid,AKIstage
         # ,AKIoffset The nephrologist suggested not to use this variable since we can not infer when AKI was developed from it. Removing it to avoid confusion.
         )
AKIlist[is.na(AKIlist)]<-0
```

### Final List of patients with AKI

```{r}
AKIlist[is.na(AKIlist)]<-0
# We add a binary version of AKI
AKIlist<-AKIlist%>%mutate(AKI_bin=if_else(AKIstage==0,0,1))
AKIlist<-AKIlist%>%filter(AKI_bin==1)%>%select(patientunitstayid)
```

## Demographics Patients

We are only interested on the first ICU stay that's why we do position
=1

```{r}
demograph_patients<-run_query(getSQL('sql/demograph_patients.sql'))
```

## Septic Patients

We are only interested on the first ICU stay that's why we do position
=1

```{r}
septic<-run_query(getSQL('sql/septic.sql'))
```

## BMI variables

```{r}
bmi_related<-run_query(getSQL('sql/bmi_related.sql'))
```

## Charlson score and comorbidities

The basic objective of this paper is to evaluate an age-comorbidity
index in a cohort of patients who were originally enrolled in a
prospective study to identify risk factors for peri-operative
complications. Two-hundred and twenty-six patients were enrolled in the
study. The participants were patients with hypertension or diabetes who
underwent elective surgery between 1982 and 1985 and who survived to
discharge. Two-hundred and eighteen patients survived until discharge.
These patients were followed for at least five years post-operatively.
The estimated relative risk of death for each comorbidity rank was 1.4
and for each decade of age was 1.4. When age and comorbidity were
modelled as a combined age-comorbidity score, the estimated relative
risk for each combined age-comorbidity unit was 1.45. Thus, the
estimated relative risk of death from an increase of one in the
comorbidity score proved approximately equal to that from an additional
decade of age. The combined age-comorbidity score may be useful in some
longitudinal studies to estimate relative risk of death from prognostic
clinical covariates. --Charlson M, Szatrowski TP, Peterson J, Gold J.
Validation of a combined comorbidity index. J Clin Epidemiol.
1994;47(11):1245-51. PMID: 7722560 --Online calculator:
--<http://www.pmidcalc.org/?sid=7722560&newtest=Y>

```{r}
charlson_score<-run_query(getSQL("sql/charlson_score.sql"))
comorbidities<-run_query(getSQL("sql/comorbidities.sql"))
```

## Mechanical Ventilation Score & bindary

Oxford Acute Severity of Illness Score (OASIS) -- The score is
calculated on the first day of each ICU patients' stay. -- Reference for
OASIS: -- Johnson, Alistair EW, Andrew A. Kramer, and Gari D. Clifford.
-- "A new severity of illness scale using a subset of acute physiology
and chronic health evaluation data elements shows comparable predictive
accuracy*." -- Critical care medicine 41, no. 7 (2013): 1711-1718. --
The score is calculated for* all\* ICU patients, with the assumption
that the user will subselect appropriate ICUSTAY_IDs. -- For example,
the score is calculated for neonates, but it is likely inappropriate to
actually use the score values for these patients.

```{r}
mechvent_score<-run_query(getSQL('sql/mechvent_score.sql'))
mechvent<-run_query(getSQL('sql/mechvent.sql'))

```

## SOFA

```{r}
sofa_total_open<-run_query(getSQL('sql/sofa/sofa_total_open.sql'))
```

## Treatment Vasopressor

```{r}
treatment_vasopressor<-run_query(getSQL('sql/treatment_vasopressor.sql'))
```

## Hypotension Time

```{r}
hypotension_time<-run_query(getSQL('sql/hypotension_time/hypotension_time.sql'))
```

## Blood gas acidosis

```{r}
bg_acidosis<-run_query(getSQL('sql/bg_acidosis.sql'))
```

## CKD history on admission

```{r}
adm_chronic_kidney_disease<-run_query(getSQL('sql/adm_chronic_kidney_disease.sql'))
```

## High resolution variables (long format)

We are going to extract them in a minute by minute rate and aggregate it
in an average every hour.

For all high resolution variables we run several aggregation steps:

1)  Median Aggregation by minute in case there are several records per
    minute per patient.

2)  Aggregation by hour by patient for the 60 minutes epoch we are
    addressing per patient going from -7 hours (-420 min) of admission
    to 24 hours (1440). i.e.:

    -   Epoch \#1 would be (-420,-360],

    -   Epoch \#2 (-360,-300] ,etc..

    -   Last window would be (1380,1440].

```{r}
labs_high_res<-run_query(getSQL('sql/high_resolution/labs_high_res.sql'))
labs_high_res$byhour <- cut(labs_high_res$chartoffset, breaks=seq(-420,1500,by = 60) ,dig.lab = 5 )
labs_high_res<-labs_high_res%>%group_by(patientunitstayid,byhour)%>%
  mutate(
    albumin = median(albumin,na.rm=T),
    bilirubin = median(bilirubin,na.rm=T),
    bun = median(bun,na.rm=T),
    calcium = median(calcium,na.rm=T),
    creatinine = median(creatinine,na.rm=T),
    glucose = median(glucose,na.rm=T),
    bicarbonate = median(bicarbonate,na.rm=T),
    totalco2 = median(totalco2,na.rm=T),
    hematocrit = median(hematocrit,na.rm=T),
    hemoglobin = median(hemoglobin,na.rm=T),
    inr = median(inr,na.rm=T),
    lactate = median(lactate,na.rm=T), # not applicable
    platelets = median(platelets,na.rm=T),
    potassium = median(potassium,na.rm=T),
    ptt = median(ptt,na.rm=T),
    sodium = median(sodium,na.rm=T),
    wbc = median(wbc,na.rm=T),
    bands = median(bands,na.rm=T),
    alt = median(alt,na.rm=T),
    ast = median(ast,na.rm=T),
    alp = median(alp,na.rm=T),
    fio2 = median(fio2,na.rm=T),
    rdw = median(rdw,na.rm=T),
    phosphate = median(phosphate,na.rm=T),
    ldh = median(ldh,na.rm=T)
        )
lab_columns<-names(labs_high_res)[names(labs_high_res) %!in% c("patientunitstayid","chartoffset","byhour")]


vitals_high_res<-run_query(getSQL('sql/high_resolution/vitals_high_res.sql'))
vitals_high_res$byhour <- cut(vitals_high_res$chartoffset, breaks=seq(-420,1500,by = 60) ,dig.lab = 5 )
vitals_high_res<-vitals_high_res%>%group_by(patientunitstayid,byhour)%>%
  summarise(
    heartrate = median(heartrate,na.rm=T),
    respiratoryrate = median(respiratoryrate,na.rm=T),
    spo2 = median(spo2,na.rm=T),
    temperature = median(temperature,na.rm=T),
    sbp = median(sbp,na.rm=T),
    dbp = median(dbp,na.rm=T),
    map = median(map,na.rm=T),
    )
vitals_columns<-names(vitals_high_res)[names(vitals_high_res) %!in% c("patientunitstayid","chartoffset","byhour")]



any_pressors_bin_high_res<-run_query(getSQL('sql/high_resolution/any_pressors_bin_high_res.sql')) 
any_pressors_bin_high_res$byhour <- cut(any_pressors_bin_high_res$chartoffset, breaks=seq(-420,1500,by = 60) ,dig.lab = 5 )
any_pressors_bin_high_res<-any_pressors_bin_high_res%>%group_by(patientunitstayid,byhour)%>%
  summarise(
    vasopressor_bin = max(vasopressor_bin,na.rm=T)
    )
any_pressors_columns<-names(any_pressors_bin_high_res)[names(any_pressors_bin_high_res) %!in% c("patientunitstayid","chartoffset","byhour")]


specific_pressors_bin_high_res<-run_query(getSQL('sql/high_resolution/specific_pressors_bin_high_res.sql'))
specific_pressors_bin_high_res[is.na(specific_pressors_bin_high_res)]<-0
specific_pressors_bin_high_res$byhour <- cut(specific_pressors_bin_high_res$chartoffset, breaks=seq(-420,1500,by = 60) ,dig.lab = 5 )
specific_pressors_bin_high_res<-specific_pressors_bin_high_res%>%group_by(patientunitstayid,byhour)%>%
  summarise(
    dopamine = max(dopamine,na.rm=T),
    dobutamine = max(dobutamine,na.rm=T),
    norepinephrine = max(norepinephrine,na.rm=T),
    phenylephrine = max(phenylephrine,na.rm=T),
    epinephrine = max(epinephrine,na.rm=T),
    vasopressin = max(vasopressin,na.rm=T),
    milrinone = max(milrinone,na.rm=T),
    heparin = max(heparin,na.rm=T)
    )
specific_pressors_columns<-names(specific_pressors_bin_high_res)[names(specific_pressors_bin_high_res) %!in% c("patientunitstayid","chartoffset","byhour")]


urine_output_high_res<-run_query(getSQL('sql/high_resolution/urine_output_high_res.sql'))
urine_output_high_res$byhour <- cut(urine_output_high_res$chartoffset, breaks=seq(-420,1500,by = 60) ,dig.lab = 5 )
urine_output_high_res<-urine_output_high_res%>%group_by(patientunitstayid,byhour)%>%
  summarise(
    outputtotal = median(outputtotal,na.rm=T),
    urineoutput = median(urineoutput,na.rm=T)
    )
urine_output_columns<-names(urine_output_high_res)[names(urine_output_high_res) %!in% c("patientunitstayid","chartoffset","byhour")]

```

## AKI ICD

```{r}
aki_icd_based <- run_query(getSQL("sql/aki/aki_icd_based.sql"))
```

## Shock ICD

```{r}
shock_icd <- run_query(getSQL("sql/shock_icd.sql"))
```

## Pressors Binary Low Resolution

```{r}
any_pressors_bin_low_res <- run_query(getSQL("sql/any_pressors_bin_low_res.sql"))
```

# Creating dataset for base model A & B

## Exclusion criteria

```{r}
print('Initial number of patients')
a<-nrow(demograph_patients)
a

print('Patients readmitted excluded:')
hr_dataset<-demograph_patients%>%filter(readmit == 0)
b<-nrow(hr_dataset)
a-b

print('Patients underage excluded:')
hr_dataset<-hr_dataset%>%filter(age_fixed >= 16)
c<-nrow(hr_dataset)
b-c

print('Patients missing mortality data excluded:')
hr_dataset<-hr_dataset%>%filter(!is.na(hospitalmortality) )
d<-nrow(hr_dataset)
c-d

# we will be excluding everyone with ESRD
# Kidney failure, also called end-stage renal disease (ESRD), is the last stage of chronic kidney disease.
print('ESRD Patients excluded:')
hr_dataset<-hr_dataset<-sqldf("SELECT A.* 
 FROM hr_dataset A
 LEFT JOIN dialysis B 
 ON A.patientunitstayid = B.patientunitstayid AND history_of_esrd_bin = 1
 WHERE B.patientunitstayid IS NULL "
 ,drv="SQLite")
e<-nrow(hr_dataset)
d-e

print('Non Septic Patients excluded:')
hr_dataset<-hr_dataset%>%filter(patientunitstayid %in% septic$patientunitstayid)
f<-nrow(hr_dataset)
e-f

print('Non Mech. Ventilated Patients excluded:')
hr_dataset<-hr_dataset<-sqldf("SELECT A.* 
 FROM hr_dataset A
 LEFT JOIN mechvent B 
 ON A.patientunitstayid = B.patientunitstayid AND mech_vent_bin = 0
 WHERE B.patientunitstayid IS NULL "
 ,drv="SQLite")
g<-nrow(hr_dataset)
f-g

# AKI is no longer an exclusion criteria, meeting 04012020
# print('Patients without AKI excluded')
# hr_dataset<-hr_dataset%>%filter(patientunitstayid %in% AKIlist$patientunitstayid)
# h<-nrow(hr_dataset)
# g-h

print('Patients with empty gender information')
hr_dataset<-hr_dataset%>%filter(gender %!in% c('','Unknown') )
h<-nrow(hr_dataset)
g-h

print('Final number of patients')
nrow(hr_dataset)

#, AKI_bin == 1 # we have decided not to use AKI as an exclusion criteria 09092020
```

## Initial JOIN (static, low resolution, variables)

```{r}
hr_dataset<-Reduce(function(...) merge(..., all.x=TRUE), list(
 hr_dataset
,bmi_related
,dialysis
,charlson_score
,mechvent_score
,sofa_total_open
,treatment_vasopressor
,hypotension_time
,bg_acidosis
,adm_chronic_kidney_disease
,mechvent
,comorbidities
,baseline_creat%>%select(patientunitstayid,creat1)
,shock_icd
,any_pressors_bin_low_res
))

# setting proper data type
hr_dataset<-unique(hr_dataset)
hr_dataset$hospitalmortality<-as.factor(hr_dataset$hospitalmortality)
hr_dataset$icd_shock<-as.factor(hr_dataset$icd_shock)
hr_dataset$vasopressor_bin_lr<-as.factor(hr_dataset$vasopressor_bin_lr)
hr_dataset$numbedscategory<-as.factor(hr_dataset$numbedscategory)
hr_dataset$dialysis_in_the_icu_bin<-as.factor(hr_dataset$dialysis_in_the_icu_bin)
hr_dataset$chf1<-as.factor(hr_dataset$chf1)

round(prop.table(table(hr_dataset$hospitalmortality)),2)
round(prop.table(table(hr_dataset$dialysis_in_the_icu_bin)),2)

round(prop.table(table(hr_dataset$hospitalmortality,hr_dataset$dialysis_in_the_icu_bin)),2)

hr_dataset_table1<-hr_dataset
```

## Static variables imputation

### Creat Imputation

We impute missing values of creat1 (0.1% missing) using median
(population) imputation.

```{r}
hr_dataset$creat1_imp<-if_else(
 is.na(hr_dataset$creat1)
,median(hr_dataset$creat1,na.rm = T)
,hr_dataset$creat1
)
```

### Drugs imputation

Binary variables: if missing, it's mean they did not receive a dose, so
we replace it with 0.

```{r}
# imputing missing binary variables, if missing, it's mean NO, Zero.
hr_dataset$vasopressor_bin_lr<-if_else(is.na(hr_dataset$vasopressor_bin_lr),'0',as.character(hr_dataset$vasopressor_bin_lr))
```

### ICD Shock imputation

Binary variables: if missing, it's mean they did not receive a dose, so
we replace it with 0.

```{r}
# imputing missing binary variables, if missing, it's mean NO, Zero.
hr_dataset$icd_shock<-if_else(is.na(hr_dataset$icd_shock),'0',as.character(hr_dataset$icd_shock))
```

### Num Beds Category imputation

Binary variables: if missing, it's mean they did not receive a dose, so
we replace it with 0.

```{r}
# imputing missing binary variables, if missing, it's mean NO, Zero.
hr_dataset$numbedscategory<-as.factor(if_else(is.na(hr_dataset$numbedscategory),'Other/Unknown',as.character(hr_dataset$numbedscategory)))
```

### Region imputation

```{r}
hr_dataset$region<-as.factor(if_else(is.na(hr_dataset$region),'Other/Unknown',as.character(hr_dataset$region)))
```

### Etnicity imputation

```{r}
hr_dataset$ethnicity<-if_else(hr_dataset$ethnicity=="",'Other/Unknown',as.character(hr_dataset$ethnicity))
```

### Apache imputation

```{r}
# -1 in apache means missing
hr_dataset$apachescore<-as.numeric(replace(hr_dataset$apachescore, hr_dataset$apachescore==-1, NA))

hr_dataset$apachescore<-if_else( is.na(hr_dataset$apachescore)
                               ,median(hr_dataset$apachescore,na.rm = T)
                               ,hr_dataset$apachescore)

```

## Data Conversion

Setting proper data type.

```{r}
hr_dataset$gender<-as.factor(hr_dataset$gender)
hr_dataset$hospitalmortality<-as.factor(hr_dataset$hospitalmortality)
hr_dataset$dialysis_in_the_icu_bin<-as.factor(hr_dataset$dialysis_in_the_icu_bin)
hr_dataset$teachingstatus<-as.factor(if_else(hr_dataset$teachingstatus==TRUE,1,0))
hr_dataset_table1$chf1<-if_else(hr_dataset_table1$chf1=='1', 'CHF','No CHF')
hr_dataset_table1$dialysis_in_the_icu_bin<-if_else(hr_dataset_table1$dialysis_in_the_icu_bin=='1', 'Dialysis','No Dialysis')
```

## Addresing Categorical Variables

### Num Beds Category

Barret Rush jue, 15 abr 17:43 (hace 13 días) para Leo, Eric, Jennifer,
Miguel

Hey guys

I was trying to code these NIS bedsizes and I dont think its really
going to work. I think we should just reclassify the eiCU ones to

Small - \<100 Medium 100-240 Large \>250

```{r}
table(hr_dataset$numbedscategory)
hr_dataset$numbedscategory<-as.character(hr_dataset$numbedscategory)
hr_dataset$numbedscategory <- plyr::revalue(hr_dataset$numbedscategory, c("<100" = "A) Small","100 - 249" = "B) Medium","250 - 499" = "C) Large", ">= 500" = "C) Large","Other/Unknown"  = "D) Other/Unknown")  )
table(hr_dataset$numbedscategory)
```

### Region

```{r}
hr_dataset$region<-relevel(as.factor(hr_dataset$region), ref= 'South')
```

### Ethnicity

04012021: We decided we are using originaal ethnicity cats

```{r}
# hr_dataset$ethnicity<-as.factor(hr_dataset$ethnicity)
# 
# # we are remaping the ethnicity
# hr_dataset$ethnicity<-if_else(hr_dataset$ethnicity %in% c('Caucasian','African American'),hr_dataset$ethnicity, as.factor('Other/Unknown') )

table(hr_dataset$ethnicity)
hr_dataset$ethnicity<-relevel(as.factor(hr_dataset$ethnicity), ref= 'Caucasian')
```

## Table 1 Base Dataset

```{r}
vars_in_base_model<-c('hospitalmortality', 'age_fixed', 'gender', 'ethnicity', 'numbedscategory', 'teachingstatus', 'region', 'dialysis_in_the_icu_bin', 'final_charlson_score', 'icd_shock')

cat_variables<-rep(NA, length(vars_in_base_model))
stratifyby<-"dialysis_in_the_icu_bin"


label(hr_dataset$age_fixed)<-'Age'

cont<-1
for (i in 1:length(vars_in_base_model) ) {
  if ( n_distinct(hr_dataset[vars_in_base_model[i] ])<=10 ) {
    print(i)
    print(vars_in_base_model[i])
    print(names(hr_dataset[vars_in_base_model[i]]))
    cat_variables[cont]<-names(hr_dataset[vars_in_base_model[i]])
    cont<-cont+1
  }
}  

cat_variables<-cat_variables[!is.na(cat_variables)]

table1_base<-print(CreateTableOne(vars = vars_in_base_model, strata = stratifyby, factorVars = cat_variables
    ,data = hr_dataset, addOverall=T),varLabels = T)

# run this in console for html output, the code below uses kableExtra::

table1_base %>%
  kbl(caption = "Table 1 of base model" , align = "c") %>%
  kable_classic_2(full_width = F, html_font = "Cambria")%>%
  add_header_above(c(" "," ", 'dialysis_in_the_icu_bin' = 2," ", "" ))
```

# Creating dataset for HR model B

## Converting static variables from wide to long format

```{r}
timeranges<-unique(labs_high_res$byhour)

hr_dataset_long<-hr_dataset%>%select(
                  patientunitstayid
                , age_fixed
                , gender
                , ethnicity
                , numbedscategory
                , teachingstatus
                , region
                , dialysis_in_the_icu_bin
                , apachescore
                , final_charlson_score
                , creat1_imp
                , icd_shock
                , vasopressor_bin_lr
                , hospitalmortality
)

# 182 entries per patient.
hr_dataset_long<-hr_dataset_long[rep(seq_len(nrow(hr_dataset_long)), each = 31), ]

# We want a chartoffset column per patient going from - 360 to 1440 

hr_dataset_long<-hr_dataset_long %>%
  group_by(patientunitstayid)%>%
  mutate(byhour = timeranges)%>%
  arrange(patientunitstayid,byhour)  

dim(hr_dataset_long)
```

## Final JOIN (high resolution variables)

```{r}
hr_dataset_long<-Reduce(function(...) merge(..., all.x=TRUE, by = c("patientunitstayid","byhour")), list(
hr_dataset_long,
labs_high_res[,names(labs_high_res)!='chartoffset'], #exclude chartoffset column
vitals_high_res[,names(vitals_high_res)!='chartoffset'], #exclude chartoffset column
any_pressors_bin_high_res,
specific_pressors_bin_high_res
))

hr_dataset_long$chartoffset<-NULL
```

## High Resolution Data Imputation

### Back-forward fill

We are using backward and forward fill for every variable by
patientunitstayid.

Continuous variables: Back and Forward fill per patient. Forward filling
means to fill missing values with previous data available for a given
patient. Backward filling means to fill missing values with the next
data point available for a given patient.

```{r}
# we don't want to apply back-forward fill to our binary columns or to our id
columns_toexclude<-c('patientunitstayid',any_pressors_columns,specific_pressors_columns)

hr_dataset_long<- hr_dataset_long %>% group_by(patientunitstayid)%>%fill(names(hr_dataset_long)[names(hr_dataset_long) %!in% columns_toexclude], .direction='downup')
```

### Binary Drugs imputation

Binary variables: if missing, it means they did not receive a dose, so
we replace it with 0.

```{r}
# imputing missing binary variables, if missing, it's mean NO, Zero.

hr_dataset_long[c('vasopressor_bin' ,'dopamine' ,'dobutamine','norepinephrine','phenylephrine','epinephrine','vasopressin','milrinone','heparin')][is.na(hr_dataset_long[c('vasopressor_bin' ,'dopamine' ,'dobutamine','norepinephrine','phenylephrine','epinephrine','vasopressin','milrinone','heparin')])] <- 0
```

## Data ordering

```{r}
hr_dataset_long<-hr_dataset_long%>%arrange(patientunitstayid,byhour)
```

# Creating dataset for HR model A

## Joining and aggregating across datasets

This is an Aggregated version of the Time Series dataset. It includes
all high resolution variables (not available in the low resolution
model) but in an aggregated fashion.

```{r}

hr_dataset_agg<-sqldf(
  'SELECT
	hr_dataset.patientunitstayid,
	hr_dataset.age_fixed,
	hr_dataset.gender,
	hr_dataset.ethnicity,
	hr_dataset.numbedscategory ,
	hr_dataset.teachingstatus,
	hr_dataset.region ,
	hr_dataset.dialysis_in_the_icu_bin,
	hr_dataset.final_charlson_score,
	hr_dataset.icd_shock,
	hr_dataset.chf1,
	hr_dataset.vasopressor_bin,
	hr_dataset.hospitalmortality,
	MEDIAN(albumin) AS albumin_agg,
	MEDIAN(bilirubin) AS bilirubin_agg,
	MEDIAN(bun) AS bun_agg,
	MEDIAN(calcium) AS calcium_agg,
	MEDIAN(creatinine) AS creatinine_agg,
	MEDIAN(glucose) AS glucose_agg,
	MEDIAN(bicarbonate) AS bicarbonate_agg,
	MEDIAN(totalco2) AS totalco2_agg,
	MEDIAN(hematocrit) AS hematocrit_agg,
	MEDIAN(hemoglobin) AS hemoglobin_agg,
	MEDIAN(inr) AS inr_agg,
	MEDIAN(lactate) AS lactate_agg,
	MEDIAN(platelets) AS platelets_agg,
	MEDIAN(potassium) AS potassium_agg,
	MEDIAN(ptt) AS ptt_agg,
	MEDIAN(sodium) AS sodium_agg,
	MEDIAN(wbc) AS wbc_agg,
	MEDIAN(bands) AS bands_agg,
	MEDIAN(alt) AS alt_agg,
	MEDIAN(ast) AS ast_agg,
	MEDIAN(alp) AS alp_agg,
	MEDIAN(fio2) AS fio2_agg,
	MEDIAN(rdw) AS rdw_agg,
	MEDIAN(heartrate) AS heartrate_agg,
	MEDIAN(respiratoryrate) AS respiratoryrate_agg,
	MEDIAN(spo2) AS spo2_agg,
	MEDIAN(sbp) AS sbp_agg,
	MEDIAN(dbp) AS dbp_agg,
	MEDIAN(map) AS map_agg,
	MEDIAN(phosphate) AS phosphate_agg,
	MEDIAN(ldh) AS ldh_agg,
	MEDIAN(temperature) AS temperature_agg,
	MAX(dopamine) AS dopamine_agg,
	MAX(dobutamine) AS dobutamine_agg,
	MAX(norepinephrine) AS norepinephrine_agg,
	MAX(phenylephrine) AS phenylephrine_agg,
	MAX(epinephrine) AS epinephrine_agg,
	MAX(vasopressin) AS vasopressin_agg,
	MAX(milrinone) AS milrinone_agg,
	MAX(heparin) AS heparin_agg
FROM
	hr_dataset
JOIN
	hr_dataset_long
USING
	(patientunitstayid)
GROUP BY
	patientunitstayid'
,drv="SQLite")

```

## Imputation

```{r}
# the following variables have more than 50% of the patients missing data, so we can't impute, therefore we are removing them

hr_dataset_agg$totalco2_agg<-NULL
hr_dataset_agg$ptt_agg<-NULL
hr_dataset_agg$bands_agg<-NULL

# median imputation for continuous columns
# most frequent value for categorical columns
impute.median <- function(x) replace(x, is.na(x), median(x, na.rm = TRUE))
impute.mostfreq <- function(x) replace(x, is.na(x), as.data.frame(sort(table(x),decreasing=TRUE))[1,1] ) 

# columns we want to median.impute contain the pattern 'agg'

hr_dataset_agg[,grep('agg',names(hr_dataset_agg))]<-as.data.frame(apply(hr_dataset_agg[,grep('agg',names(hr_dataset_agg))],2, function(x) impute.median(x) ))

hr_dataset_agg$vasopressor_bin<-impute.mostfreq(hr_dataset_agg$vasopressor_bin)

```

## Table 1 High Resolution Dataset model A

```{r}
vars_in_hr_model<-c("age_fixed","gender","ethnicity","numbedscategory","teachingstatus","region","dialysis_in_the_icu_bin","final_charlson_score","icd_shock","chf1","vasopressor_bin","hospitalmortality","albumin_agg","bilirubin_agg","bun_agg","calcium_agg","creatinine_agg","glucose_agg","bicarbonate_agg","hematocrit_agg","hemoglobin_agg","inr_agg","lactate_agg","platelets_agg","potassium_agg","sodium_agg","wbc_agg","alt_agg","ast_agg","alp_agg","fio2_agg","rdw_agg","heartrate_agg","respiratoryrate_agg","spo2_agg","sbp_agg","dbp_agg","map_agg","phosphate_agg","ldh_agg","temperature_agg","dopamine_agg","dobutamine_agg","norepinephrine_agg","phenylephrine_agg","epinephrine_agg","vasopressin_agg","milrinone_agg","heparin_agg")

cat_variables<-rep(NA, length(vars_in_hr_model))
stratifyby<-"dialysis_in_the_icu_bin"


label(hr_dataset_agg$age_fixed)<-'Age'

cont<-1
for (i in 1:length(vars_in_hr_model) ) {
  if ( n_distinct(hr_dataset_agg[vars_in_hr_model[i] ])<=10 ) {
    print(i)
    print(vars_in_hr_model[i])
    print(names(hr_dataset_agg[vars_in_hr_model[i]]))
    cat_variables[cont]<-names(hr_dataset_agg[vars_in_hr_model[i]])
    cont<-cont+1
  }
}  

cat_variables<-cat_variables[!is.na(cat_variables)]

table1_base<-print(CreateTableOne(vars = vars_in_hr_model, strata = stratifyby, factorVars = cat_variables
    ,data = hr_dataset_agg, addOverall=T),varLabels = T)

# run this in console for html output, the code below uses kableExtra::

table1_base %>%
  kbl(caption = "Table 1 of base model" , align = "c") %>%
  kable_classic_2(full_width = F, html_font = "Cambria")%>%
  add_header_above(c(" "," ", 'dialysis_in_the_icu_bin' = 2," ", "" ))
```
