---
title: "02_model_training"
author: "Miguel Ángel Armengol"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_notebook:
    code_folding: hide
    number_sections: yes
    theme: flatly
    toc: yes
    toc_float: yes

knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = paste0(substr(inputFile,1,nchar(inputFile)-4)," ",Sys.Date(),'.html')) })
---

# Libraries

```{r message=FALSE, warning=FALSE}
library(summarytools)
library(RPostgreSQL)
library(dplyr)
library(sqldf)
#library(MatchIt)
library(ggplot2)
library(survey)
library(tableone)
library(sandwich)
library(mgcv)
library(lme4)
library(table1)
```

# Low resolution model

```{r}
print('Patients in the dataset')
n_distinct(hr_dataset$patientunitstayid)

#checking there is not missingness in any variable
table1( ~ hospitalmortality + age_fixed + gender + ethnicity + numbedscategory + teachingstatus + region + dialysis_in_the_icu_bin + apachescore + final_charlson_score + creat1_imp + icd_shock + vasopressor_bin_lr, data = hr_dataset)

low_res_model<-glm(hospitalmortality ~ 
                  age_fixed
                + gender
                + ethnicity
                
                + dialysis_in_the_icu_bin
                + apachescore
                + final_charlson_score
                + creat1_imp
                + icd_shock
                + vasopressor_bin_lr
                , family=binomial
                , data=hr_dataset)
print('Patients included in the model')
nobs(low_res_model)

OR_table<-as.data.frame(round(exp(cbind(OR=coef(low_res_model), confint.default(low_res_model))),2))

OR_table
```

# Creating our dataset

```{r}
# we need a dataset with no missing values so we are subsetting the original one so it includes only the required variables for the Propensity score matching which actually never have missing values.

hr_prop_score_dataset<-hr_dataset_long%>%dplyr::select(
  patientunitstayid
, age_fixed
, gender
, ethnicity
, numbedscategory
, teachingstatus
, region
, dialysis_in_the_icu_bin
, apachescore
, final_charlson_score
, creat1_imp
, icd_shock
, vasopressor_bin_lr
, hospitalmortality
, albumin
, bilirubin
, bun
, calcium
, creatinine
, glucose
, bicarbonate
, totalco2
, hematocrit
, hemoglobin
, inr
, lactate
, platelets
, potassium
, ptt
, sodium
, wbc
, bands
, alt
, ast
, alp
, fio2
, rdw
, heartrate
, respiratoryrate
, spo2
, sbp
, dbp
, map
, temperature
, vasopressor_bin
, dopamine
, dobutamine
, norepinephrine
, phenylephrine
, epinephrine
, vasopressin
#, milrinone # milrinone is almost completely missing and is confusing the model
, heparin
)

saveRDS(hr_prop_score_dataset,"hr_prop_score_dataset.rds")
```

# Summary

```{r}
df_for_summary<-as.data.frame(hr_prop_score_dataset)
summarytools::view(dfSummary(df_for_summary))
```

# Random effect model

## Data wrangling before modeling

Some predictor variables are on very different scales: so we should rescale.

### Log transforming

We are log transforming

```{r}
hr_prop_score_dataset[,c("age_fixed" , "apachescore" , "final_charlson_score" , "creat1_imp" , "albumin" , "bilirubin" , "bun" , "calcium" , "creatinine" , "glucose" , "bicarbonate" , "totalco2" , "hematocrit" , "hemoglobin" , "inr" , "lactate" , "platelets" , "potassium" , "ptt" , "sodium" , "wbc" , "bands" , "alt" , "ast" , "alp" , "fio2" , "rdw" , "heartrate" , "respiratoryrate" , "spo2" , "sbp" , "dbp" , "map" , "temperature")]<-log(hr_prop_score_dataset[,c("age_fixed" , "apachescore" , "final_charlson_score" , "creat1_imp" , "albumin" , "bilirubin" , "bun" , "calcium" , "creatinine" , "glucose" , "bicarbonate" , "totalco2" , "hematocrit" , "hemoglobin" , "inr" , "lactate" , "platelets" , "potassium" , "ptt" , "sodium" , "wbc" , "bands" , "alt" , "ast" , "alp" , "fio2" , "rdw" , "heartrate" , "respiratoryrate" , "spo2" , "sbp" , "dbp" , "map" , "temperature")])
```

### Mean centering and standardizing

We are re-scaling and and mean centering all mets. After this process is complete they all should have a mean of 0 and an SD of 1.

```{r}
scale_centered_scaled<-function(x){
  scaled_centered <- scale(x, center=T, scale=T)
  return(scaled_centered)
}

hr_prop_score_dataset[,c("age_fixed" , "apachescore" , "final_charlson_score" , "creat1_imp" , "albumin" , "bilirubin" , "bun" , "calcium" , "creatinine" , "glucose" , "bicarbonate" , "totalco2" , "hematocrit" , "hemoglobin" , "inr" , "lactate" , "platelets" , "potassium" , "ptt" , "sodium" , "wbc" , "bands" , "alt" , "ast" , "alp" , "fio2" , "rdw" , "heartrate" , "respiratoryrate" , "spo2" , "sbp" , "dbp" , "map" , "temperature")]<-scale_centered(hr_prop_score_dataset[,c("age_fixed" , "apachescore" , "final_charlson_score" , "creat1_imp" , "albumin" , "bilirubin" , "bun" , "calcium" , "creatinine" , "glucose" , "bicarbonate" , "totalco2" , "hematocrit" , "hemoglobin" , "inr" , "lactate" , "platelets" , "potassium" , "ptt" , "sodium" , "wbc" , "bands" , "alt" , "ast" , "alp" , "fio2" , "rdw" , "heartrate" , "respiratoryrate" , "spo2" , "sbp" , "dbp" , "map" , "temperature")])

```

# Base Model A

Low resolution (wide format).

```{r}
log_reg <- glm( hospitalmortality ~ 
                  age_fixed
                + gender
                + ethnicity 
                + dialysis_in_the_icu_bin
                + final_charlson_score
                + icd_shock
                + chf1
                ,family = 'binomial'
                , data=hr_dataset)

summary(log_reg)

OR_table<-as.data.frame(round(exp(cbind(OR=coef(log_reg), confint.default(log_reg))),2))

OR_table
```

# Base Model B

Low resolution (wide format).

```{r}
log_reg <- glm(   hospitalmortality ~ 
                  age_fixed
                + gender
                + ethnicity
                + numbedscategory 
                + teachingstatus
                + region 
                + dialysis_in_the_icu_bin
                + final_charlson_score
                + icd_shock
                + chf1
                , family='binomial'
                , data=hr_dataset)

summary(log_reg)

OR_table<-as.data.frame(round(exp(cbind(OR=coef(log_reg), confint.default(log_reg))),2))

OR_table
```

# HR Model A

Addresses all high resolution variables (not available in the low resolution model) but in an aggregated fashion.

```{r}
log_reg <- glm( hospitalmortality ~ . - patientunitstayid 
                , family='binomial'
                , data=hr_dataset_agg)

summary(log_reg)

OR_table<-as.data.frame(round(exp(cbind(OR=coef(log_reg), confint.default(log_reg))),2))

OR_table
```

# HR Model B

Time Series.

**Not for this project.**

# Causal inference

Ref.: <https://www.youtube.com/watch?v=P2LArUuzmso>

## Propensity Score - logistic regression

```{r}
# Fit a propensity score model: logistic regression
ps_model <- glm(dialysis_in_the_icu_bin ~ 
                  apachescore
                + age_fixed
                + gender
                + final_charlson_score
                + ethnicity
                + creat1_imp
                , family=binomial()
                , data=hr_prop_score_dataset)

summary(ps_model)

OR_table<-as.data.frame(round(exp(cbind(OR=coef(ps_model), confint.default(ps_model))),2))

OR_table

# Value of propensity score for each subject

pscore<-ps_model$fitted.values

treatment_pscore<-cbind(hr_prop_score_dataset%>%select(dialysis_in_the_icu_bin),pscore)

ggplot(treatment_pscore, aes(x=pscore, fill = dialysis_in_the_icu_bin)) + 
  geom_histogram(alpha = 0.5,position="identity") +
  scale_fill_manual(name="RRT",values=c("#1abc9c","#f1c40f"),labels=c("No (Control)","Yes (Treated)"))+theme_minimal()
```

## Inverse probability weighting

```{r}
# create weight
hr_prop_score_dataset<-hr_prop_score_dataset%>%mutate(
  weight=if_else(dialysis_in_the_icu_bin==1
                 ,1/(pscore)
                 ,1/(1-pscore))
)

ggplot(hr_prop_score_dataset, aes(x = weight, fill = dialysis_in_the_icu_bin)) +
   geom_density(alpha = 0.5, colour = "grey50") +
   geom_rug() +
   scale_x_log10(breaks = c(1, 5, 10, 20, 40)) +
   ggtitle("Distribution of inverse probability weights")+ scale_fill_manual(name="RRT",values=c("#1abc9c","#f1c40f"),labels=c("No (Control)","Yes (Treated)"))+theme_minimal()

# apply weights to data
weighted_data<-svydesign(ids = ~ patientunitstayid, data = hr_prop_score_dataset,
                         weights = ~weight
                         )

# weighted table 1

weighedtable<- svyCreateTableOne( vars= c("apachescore", "age_fixed","gender","final_charlson_score","ethnicity","creat1_imp"), strata = "dialysis_in_the_icu_bin", data = weighted_data, test = F)

# weighted table 1 wih standarized mean differences (SMD)

as.data.frame(print(weighedtable,smd=T))
```

In tableone: Ignore SD and sample sizes, ONLY SMD is reliable

## Marginal structural modeling

### Relative risk with CI

```{r warning=FALSE}
# Obtaining causal relative risk. Weighed GLM

glm_model_rr<-glm(
  hospitalmortality ~ 
  dialysis_in_the_icu_bin # treatment is the only exposure now
, weights = weight  
, family = binomial(link = log) # we are using the log link since we are interested in the relative risk
, data= hr_dataset
)

# summary of the glm_model_rr final model
# beta Inverse probability weighting
betaiptw<-coef(glm_model_rr)

# to properly account for weighting, we are going to use asymptotic (sandwich) variance

SE<-sqrt(diag(vcovHC(glm_model_rr,type = "HC0"))) # getting the standard error.

# we get point estimate and CI for relative risk

beta_causal_relative_risk<-exp(betaiptw[2]) # we need to exponientiate since we logged before
lCI<-exp(betaiptw[2]-1.96*SE[2])
uCI<-exp(betaiptw[2]+1.96*SE[2])

as.data.frame(cbind(beta_causal_relative_risk, lCI, uCI))
```

### Risk difference with CI

```{r warning=FALSE}
# Obtaining Obtaining risk difference with CI

glm_model_diff<-glm(
  hospitalmortality ~ 
  dialysis_in_the_icu_bin # treatment is the only exposure now
, weights = weight  
, family = binomial(link = 'identity') # we are using the identity link since we are interested in isk difference
, data= hr_dataset
)

# summary of the glm_model_diff final model
# beta Inverse probability weighting
betaiptw<-coef(glm_model_diff)

# to properly account for weighting, we are going to use asymptotic (sandwich) variance

SE<-sqrt(diag(vcovHC(glm_model_diff,type = "HC0"))) # getting the standard error.

# we get point estimate and CI for relative risk

beta_risk_difference<-exp(betaiptw[2]) # we need to exponientiate since we logged before
lCI<-exp(betaiptw[2]-1.96*SE[2])
uCI<-exp(betaiptw[2]+1.96*SE[2])

as.data.frame(cbind(beta_risk_difference, lCI, uCI))
```
